Leadership across cultures and contexts: an exploratory analysis

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Minor changes in abstract and conclusion.

Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their large-scale, complex organizations. Many researchers have turned to evolutionary theory to better understand this diversity. Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

@@

Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face. The diversity within the non-WEIRD, however, should not be overlooked. Leadership theories, including evolutionary and managerial approaches have not previously had broad systematic, cross-cultural data on leadership, management, and social organization across diverse populations. A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.

@@

<!-- CLOSING STATEMENT -->
Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics. Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].

>>>>

Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their large-scale, complex organizations. Many researchers have turned to evolutionary theory to better understand this diversity. Evolutionary theories of leadership, in turn, draw on ethnographic cases of more traditional societies thought to more closely resemble the smaller-scale, face-to-face societies in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

@@

Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence crosses domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face. The diversity within the non-WEIRD, however, should not be overlooked. Leadership theories, including evolutionary and managerial approaches have not previously had broad systematic, cross-cultural data on leadership, management, and social organization across diverse populations. A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.

@@

<!-- CLOSING STATEMENT -->
Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics. Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @smith_cultural_2020; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mattison_evolution_2016].
